---
title: "LifeCycleSavings - CUONG VO - 131116"
output:
  pdf_document: default
  html_document: default
date: "2023-11-14"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Initial setup

```{r echo=TRUE}
library(dplyr)
attach(LifeCycleSavings)
```

## Dataset Desciption

Dataset is about the savings of personal based on disposable income in the 10 year period from 1960 to 1970

```{r echo=TRUE, paged.print=TRUE}
summary(LifeCycleSavings)
```

```{r echo=TRUE, paged.print=TRUE}
str(LifeCycleSavings)

```

The dataset is in the structure of dataframe, with 5 columns about ratio of savings (divided by disposalbe income) ("sr"), % of population ("pop15", "pop75"), disposable income ("dpi") and change of disposable income ("ddpi")

```{r echo=TRUE}
mean(sr)
sd(sr)
```

For the savings ratio, the average amount in the population is `r mean(sr)` and the standard deviation is `r sd(sr)` and it's told that, people tend to save around `r mean(sr)`% their disposal income.

```{r echo=TRUE}
cor(LifeCycleSavings[, c("sr","pop15", "pop75", "dpi", "ddpi")])
```
```{r echo=TRUE}
boxplot(sr, xlab = 'Savings Ratio')
boxplot(pop15, xlab = '% population under 15')
boxplot(pop75, xlab = '% population over 75')
boxplot(dpi, xlab = 'Disposable income per capita')
boxplot(ddpi, xlab = 'Variance of DPI')
```

The dataset do not have any outlier, base on the boxplot we can define for the outlier, the data do not have the outlier.

```{r echo=TRUE}
hist(sr, main = 'Histogram of Savings Ratio', probability = TRUE)
lines(density(sr, adjust = 1.5), col = 'red', lwd = 2)
```
Base on the histogram, the histogram of the savings ratio, main the observation places from 0-15%

## Top N Analysis

```{r echo = TRUE}
top10sr <- arrange(LifeCycleSavings, desc(sr)) %>% select('sr')
top10sr<- top10sr %>% top_n(10)
print(top10sr)
```

```{r echo = TRUE}
barplot(top10sr$sr, 
        names = row.names(top10sr), 
        xlab = 'Country', 
        ylab = 'Percent', 
        col = if (top10sr$sr[1]) {'red'})
```

```{r echo = TRUE}
top10dpi <- arrange(LifeCycleSavings, desc(dpi)) %>% select('dpi')
top10dpi<- top10dpi %>% top_n(10)
print(top10dpi)
```

```{r echo = TRUE}
barplot(top10dpi$dpi, 
        names = row.names(top10dpi), 
        xlab = 'Country', 
        ylab = 'Percent', 
        col = if (top10dpi$dpi[1]) {'red'})
```

As we can see *`r row.names(top10sr)[1]`* have the highest rate of savings in the population at *`r top10sr$sr[1]`%* but it seem like *`r row.names(top10sr)[1]`* is not the country have the highest disposable income per capita. Compared with the country which have the highest dpi, *`r row.names(top10dpi)[1]`* can not make it into top 10 country that have the highest savings rate. 

## Correlation Analysis

```{r}
require(stats); require(graphics)
pairs(LifeCycleSavings, panel = panel.smooth,
      main = "LifeCycleSavings data")
```

As we can see from correlation matrix, the saving ratio and % population under 15 is have negative correlation. Countries with young population tend to spend more instead of savings. However, the countries have old population tend to have higher savings ratio.